Stack Safety as a Security Property

### 2020-07-08 main.tex
More formal stack testing, checkpoint

\definition{Variant stacks}
  The testing property keeps track of relevant information on a stack whose
  elements represent variants of the machine by recording an initial machine
  state, an initial and current variant of the machine, a contour and a return
  map:
%
  $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS \times \RETMAPS$$
  $$\vse = (\mathit{init\_machine}(\vse), \mathit{init\_variant}(\vse), \mathit{curr\_variant}(\vse), \contour(\vse), \mathit{ret}(\vse))$$
  $$\vs \in \VSS$$
%
  Given a machine state $\mach$, a stack $\vs$ is \emph{well-formed} with
  respect to it if the stack is nonempty, the return map of the bottom element
  is the false property, and for each element $\vse \in \vs$:
  \begin{itemize}
  \item $\mathit{curr\_variant}(\vse)$ is a variant of $\mach$.
  \item $\mach$ is in the trace $\machT$ of its initial machine,
    $\mathit{init\_machine}(\vse) \hookrightarrow \machT$.
  \item Its variant state $\mathit{curr\_variant}(\vse)$
    is in the trace of its initial variant,
    $\mathit{init\_variant}(\vse) \hookrightarrow \machT$.
  \item The return function is in the return map if not the last element. \rb{TODO}
  \item The registers of its contour $\contour(\vse)$ are public.
  \end{itemize}

@@

\definition{Eager testing property}
  Let $\callmap$ a call map and $\codemap$ a code map. Given a state
  $(\mach, \pol) \in \MPS$ that is well-formed w.r.t. a variant stack $\vs$, we
  say that an MP-trace induced from $(\mach, \pol)$ satisfies the \emph{eager
  testing property} at a given call depth if:

@@

  \item If the machine fail-stops, i.e.,
    $(\mach, \pol) \hookrightarrow_P (\mach, \pol, \tau)$,
    the property is safisfied for the singleton trace $[(\mach, \pol)]$ at any
    depth up to the size of $\vs$.

  \item If the top variant of the stack (and only the top variant) returns,
    the property is safisfied for the singleton trace $[(\mach, \pol)]$ at depth
    equal to the size of the stack. \rb{Avoid confusion with return below}

  \item If (1) the machine takes a step,
    $(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)$,
    that is neither a call nor a return, (2) all
    variants can also take a step $\vs \hookrightarrow \vs'$ \rb{notation for this}
    while satisfying the eager integrity and confidentiality tests \rb{more formal?}
    (these tests are required also for calls and returns, and omitted in those cases),
    and (3) the post-step state $(\mach', \pol')$ and stack $\vs'$ satisfy the eager testing property
    for a given depth $d$ \rb{omitting stack size condition} and MP-trace $\MPT$,
    then the property is also satisfied for trace $(\mach, \pol) : \MPT$
    at depth $d$.

  \item If the machine takes a call step
    $(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)$,
    the contour $\contour$ is updated based on the top of the current machine.
    For the property to hold for a trace $(\mach, \pol) \cdot \MPT$ at depth $d$
    (and in addition to the general step conditions),
    the eager integrity test must hold for $\contour$ at the call,
    the eager confidentiality test must hold for all variants at the call,
    including the new variant that is added to the top of the stack,
    and the property recurses on an arbitrary variant at the new contour
    at depth $d$ with trace $\MPT$.

  \item If the machine takes a return step at the top of the variant stack
    $(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)$,
    the property holds for a trace $(\mach, \pol) \cdot \MPT$ at depth $d$
    if (in addition to the general step conditions)
    the property recurses at depth $d$ with trace $\MPT$
    after removing the returning top variant from the stack.

  \item If the machine takes a return step that is also a call, the last two
    cases are combined.

@@

>>>>

\definition{Variant stacks}
  The testing property keeps track of relevant information on a stack whose
  elements represent variants of the machine by recording an initial machine
  state, an initial and current variant of the machine, a contour and a return
  map:
%
  $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS \times \RETMAPS$$
  $$\vse = (\mathit{init\_machine}(\vse), \mathit{init\_variant}(\vse), \mathit{curr\_variant}(\vse), \contour(\vse), \mathit{ret}(\vse))$$
  $$\vs \in \VSS$$
%
  Given a machine state $\mach$, a stack $\vs$ is \emph{well-formed} with
  respect to it if the stack is nonempty, the return map of the bottom element
  is the false property, and for each element $\vse \in \vs$:
  \begin{itemize}
  \item $\mathit{curr\_variant}(\vse)$ is a variant of $\mach$.
  \item $\mach$ is in the trace $\machT$ of its initial machine,
    $\mathit{init\_machine}(\vse) \hookrightarrow \machT$.
  \item Its variant state $\mathit{curr\_variant}(\vse)$
    is in the trace of its initial variant,
    $\mathit{init\_variant}(\vse) \hookrightarrow \machT$.
  \item The return function is in the return map if not the last element. \rb{TODO}
  \item The registers of its contour $\contour(\vse)$ are public.
  \end{itemize}
%
  A variant stack steps if \rb{\ldots, explain, fix notation}, $\vs
  \hookrightarrow \vs'$.

@@

\definition{Coinductive stack testing property}
%
  Let $\callmap$ a call map and $\codemap$ a code map. Given a state $(\mach,
  \pol) \in \MPS$ that is well-formed w.r.t. a variant stack $\vs$ and induces
  an MP-trace $(\mach, \pol) \hookrightarrow \MPT$, we say that the \emph{stack
  testing property} is satisfied at call depth $d$, and write
  $\mathit{stack}_T ~ d ~ \MPT ~ \vs$, as follows:

@@

  \item If the machine \emph{fail-stops}, i.e., $(\mach, \pol) \hookrightarrow_P
    (\mach, \pol, \tau)$, the property is safisfied for the singleton trace:
    $\mathit{stack}_T ~ d ~ [(\mach, \pol)] ~ \vs$, for any $d$ up to the size of
    $\vs$. (That is, a violation was detected $d$ levels into the stack.
    Elements are created and pushed to the variant stack on calls and popped
    from it on returns, so its size always reflects the depth of the program
    stack.)

  \item If the top variant of the stack (and only the top variant) returns,
    i.e., $\mathit{ret}(\head(\vs), \mach)$ holds, the property is safisfied for
    the singleton trace $[(\mach, \pol)]$ at depth equal to the size of the
    stack: $\mathit{stack}_T ~ |\vs| ~ [(\mach, \pol)] ~ \vs$. \rb{Explain the
    use of this vs the return case below.}

  \item If
    %
    (1) the machine takes a \emph{non-call, non-return step} $(\mach,
    \pol) \hookrightarrow_P (\mach', \pol', \obs)$ and the variant stack steps
    to $\vs \hookrightarrow \vs'$,
    %
    (2) the variant stack test $\mathit{\vs}_T ~
    (\mach, \pol) ~ \vs$ is satisfied, and
    %
    (3) the stack safety test
    $\mathit{stack}_T ~ d ~ \MPT ~ \vs'$ is coinductively satisfied at depth $d$
    and for a trace starting continuing after the step, $\head(\MPT) = (\mach',
    \pol')$,
    %
    {\em then} we conclude $\mathit{stack}_T d ~ ((\mach, \pol) : \MPT) ~ vs$.
    \rb{omitting stack size condition}

  \item
%
    If (1) the machine takes a \emph{call step} $(\mach, \pol) \hookrightarrow_P
    (\mach', \pol', \obs)$ and the variant stack steps to $\vs \hookrightarrow
    \vs'$,
%
    (2) $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$,
%
    (3) for the new contour at the call, $\contour$, \rb{definition?} the
    machine satisfies the stack integrity test $\mathit{int}_T ~ \contour ~
    \mach ~ \mach'$ and the stack confidentiality test with regard to any
    variant $\mathit{variant} ~ C ~ \mach ~ \mach_{\mathit{var}}$,
    $\mathit{conf}_T ~ \mach ~ \mach_{\mathit{var}}$ \rb{definition?},
%
    and (4) the stack safety test is coinductively satisfied at depth $d$ for a
    continuing trace $\MPT$ where $\head(\MPT) = (\mach', \pol')$ after an
    arbitrary variant machine derived from $\contour$ has been added to the top
    of the stack, $\mathit{stack}_T ~ d ~ \MPT ~ (\vse : \vs')$, \rb{define
      $\vse$}
%
    {\em then} we conclude $\mathit{stack}_T ~ d ~ ((\mach, \pol) : \MPT) ~ vs$.

    %% the contour $\contour$ is updated based on the top of the current machine.
    %% For the property to hold for a trace $(\mach, \pol) \cdot \MPT$ at depth $d$
    %% the eager integrity test must hold for $\contour$ at the call,
    %% the eager confidentiality test must hold for all variants at the call,
    %% including the new variant that is added to the top of the stack,
    %% and the property recurses on an arbitrary variant at the new contour
    %% at depth $d$ with trace $\MPT$.

  \item If the machine takes a \emph{return step} at the top of the variant stack
    $(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)$,
    the property holds for a trace $(\mach, \pol) \cdot \MPT$ at depth $d$
    if (in addition to the general step conditions)
    the property recurses at depth $d$ with trace $\MPT$
    after removing the returning top variant from the stack.

  \item If the machine takes \emph{simultaneous return-call step}, that is, the
    instruction to which execution returns immediately after completing a call
    is itself a call instruction, the last two cases are combined.

@@

\definition{Top-level stack testing property}
%
  \ldots wrap the coinductive stack testing property
